---
title: "finding-naughty-non-profits-cities"
author: "Alexander Nielson"
date: "3/10/2020"
output: html_document
---
I messed up in the last report and only did Counties! So this will fix to include all the different types in cities. 

# Libraries 
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(odbc)

library(fuzzyjoin)
```

```{r}
dsn_aws        <- "transpAWS"
dsn_salesforce <- "salesforce"

odbc_aws <- dbConnect(odbc::odbc(), dsn_aws)
odbc_sf  <- dbConnect(odbc::odbc(), dsn_salesforce)

rm(dsn_aws, dsn_salesforce)
```

##load Lookup table for uca
```{r}
osa_lookup_file_name <- "osa_lookup.xlsx"

osa_lookup <- excel_sheets(osa_lookup_file_name) %>%
  map(read_excel, path = osa_lookup_file_name)

names(osa_lookup) <- c("fund", "funct", "account")

rm(osa_lookup_file_name)


```

```{r}
grant_accounts <- osa_lookup %>%
  pluck("account") %>%
  filter(str_detect(description %>% str_to_lower(),"grant"))
grant_accounts 
```


```{r}

# go to secure.utah.gov and get the batch_id.  
 
report<- 
      dbGetQuery(
        odbc_aws,
        paste("
              SELECT 
                e.name AS entity_name,
                e.id AS entity_id,
                e.govt_lvl AS entity_gov_level,
                
                t.amount AS amount,
                t.type AS type,
                t.account_number AS account_number,
                t.fiscal_year AS fiscal_year,
                
                v.name AS vendor_name,
                v.id AS vendor_id
              FROM entity e
              JOIN batch b ON b.entity_id = e.id
              JOIN transaction t ON b.id = t.batch_id
              JOIN vendor v ON v.id = t.vendor_id
              WHERE govt_lvl = 'CITY'
              AND (t.fiscal_year = '2018' OR t.fiscal_year = '2019')

              "))

#2,757,462 records queired. 
```

```{r}
report <- report %>% 
    filter(str_detect(
      .$account_number,
      regex("^[:digit:]{3}-[:digit:]{6}-[:digit:]{8}$")
    )) %>% 
  mutate(
     account = .[["account_number"]] %>%  substr(12, 19)
  )



transactions_with_grants<- report %>% 
  inner_join(grant_accounts, by = c("account"="number"))

transactions_with_grants_no_name_provided <- transactions_with_grants %>% filter(vendor_name =="Not Applicable")

transactions_with_grants <- transactions_with_grants %>% filter(vendor_name !="Not Applicable")

```

so there are 65,508 transactions with 

```{r}
transactions_with_grants_rev <- transactions_with_grants %>% 
  filter(type==2)

transactions_with_grants_exp <- transactions_with_grants %>% 
  filter(type==1)
```


```{r}
exp_agg<- transactions_with_grants_exp %>% group_by(vendor_id, vendor_name, fiscal_year) %>% 
  summarize(net_amount = sum(amount),
            num_transactions = n())
```

```{r}
exp_agg_g25k<- exp_agg %>% filter(net_amount >=25000)
```


#Query Salesforce data
```{r}
sf_non_profits <- 
  dbGetQuery(
    odbc_sf,
   
    "SELECT 
      a.Name                          AS name,
      a.Id                            AS id,
      a.Transparency_ID__c            AS t_id,
      a.Fiscal_Year_Begins__c         AS begin_fy,
      a.Expense_Revenue_Start_Date__c AS begin_report_er, 
      a.Wage_Start_Date__c            AS begin_report_w2,
      r.Name AS govt_type
    FROM Account         AS a
    LEFT JOIN RecordType AS r
      ON a.RecordTypeId = r.Id
    WHERE a.RecordTypeId IN (
      SELECT Id
      FROM RecordType
      WHERE SobjectType = 'Account'
      AND Name  IN ('Non Profits'))
    AND a.Name NOT IN (
      'Intermountain Power Agency',
      'test city 2',
      'Utah Associated Municipal Power Systems',
      'Utah Municipal Power Agency')
    --AND Entity_Status__c NOT IN ('Inactive', 'Dissolved')
    
    ") %>% 
  as_tibble() %>% 
  mutate(
  govt_type = 
    if_else(
      name %in% 
        c("State of Utah",
          "Utah System of Higher Education - Student Loan Guarantee Program",
          "Utah System of Higher Education - Student Loan Purchase Program",
          "Utah System of Higher Education - Utah Educational Savings Plan dba my529",
          "Utah System of Higher Education - Utah State Board of Regents"),
      "Monitored State Agency",
      govt_type)) %>% 
  filter(govt_type != "State of Utah (agencies/depts/comp units/ etc.)") %>% 
  # The State of Utah is required to report to Transparent Utah, but in most instances we do not monitor its data:
  filter(name != "State of Utah") #%>% 
  #select(-govt_type)

```

go make sure that the number of non profits matches the salesforce report that Calvin created. 

lets make all the names loewr case to try and match easier
```{r}
# exp_agg_g25k <- exp_agg_g25k %>% 
#   ungroup() %>% 
#   mutate(vendor_name = str_to_lower(vendor_name))
# 
# sf_non_profits <- sf_non_profits %>% 
#   mutate(name = str_to_lower(name))
# 
# # now I need to get rid of all the inc and corp stuff. 
# 
# exp_agg_g25k <- exp_agg_g25k %>% 
#   mutate(vendor_name = str_replace_all(vendor_name,"inc|corp","" ),
#          vendor_name = str_replace_all(vendor_name,"[:punct:]","" ))
# 
# sf_non_profits <- sf_non_profits %>% 
#   mutate(name = str_replace_all(name,"inc|corp","" ),
#          name = str_replace_all(name,"[:punct:]","" ))

```



```{r}
# temp_join <- exp_agg_g25k %>% anti_join(sf_non_profits, by = c("vendor_name" = "name"))
```


```{r}
# temp<- lapply(exp_agg_g25k$vendor_name, agrep, sf_non_profits$name, value = TRUE)
```


```{r}
twg_over_25k <- transactions_with_grants %>% group_by(vendor_id, vendor_name,entity_name, entity_id, fiscal_year, account, description, level_primary, level_secondary, level_tertiary, level_quaternary) %>% 
  summarize(net_amount = sum(amount),
            num_twg = n()) %>% 
  filter(net_amount>25000)

twg_over_25k <- twg_over_25k %>%
  ungroup() %>% 
  mutate(vendor_name = str_to_lower(vendor_name)) %>%
  filter(vendor_name != "not provided" & vendor_name !="not applicable") %>% 
  mutate(
   # vendor_name = str_replace_all(vendor_name, "inc|corp|corporation|incorporation", ""),
    vendor_name = str_replace_all(vendor_name, "[:punct:]", "")
  )

sf_non_profits <- sf_non_profits %>%
  mutate(name = str_to_lower(name)) %>%
  mutate(#name = str_replace_all(name, "inc|corp|corporation|incorporation", ""),
         name = str_replace_all(name, "[:punct:]", ""))

```


```{r}
test_fuzzy_join<- twg_over_25k %>% 
  stringdist_inner_join(sf_non_profits, by = c("vendor_name" = "name"), max_dist = 3)
#manually checked that all were acceptable.
```

```{r}
test_fuzzy_join %>% select(vendor_name, name)
```


```{r}
red_twg_over_25k <- twg_over_25k %>% anti_join(test_fuzzy_join, by=c("vendor_id" = "vendor_id"))
```

lets get the names of all cities so we can remove them. 

```{r}
cities<- 
      dbGetQuery(
        odbc_aws,
        paste("
              SELECT name
              FROM entity
              WHERE govt_lvl = 'CITY'
        
              "))
```

```{r}
red_twg_over_25k %>% filter(!str_detect(vendor_name, "city"))%>%  pull(vendor_name) %>% unique()
```


```{r}
write_csv(red_twg_over_25k,"E:/rprojects/Transparency-Ad-Hoc/2020-03-03-finding-naughty-non-profits/new-non-profits-cities-only.csv")
```

